---
title: "Assignment 1 - Loading Data into a Data Frame"
author: "Jayden Jiang"
date: "2025-02-02"
output: html_document
---

## Library
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Introduction
The article I chose from FiveThirtyEight was "The Ultimate Halloween Candy Power Ranking," published on October 31, 2016. The article presents a ranking based on survey, sales data, and other metrics, providing insights into which treats are most favored during the Halloween season. 

The dataset was taken from: https://fivethirtyeight.com/videos/the-ultimate-halloween-candy-power-ranking/

## Load Data
```{r}
candy_rawdata <- read.csv("https://raw.githubusercontent.com/JaydeeJan/Best-Candy/refs/heads/main/candy-data.csv")
head(candy_rawdata)
```

### Columns Renaming & Subsetting
```{r}
# rename 
candy_rawdata <- candy_rawdata %>%
  rename(
    Candy_Name = competitorname,
    Chocolate = chocolate,
    Fruity = fruity,
    Caramel = caramel,
    Peanut_Almond = peanutyalmondy,
    Nougat = nougat,
    Crisped_Rice_Wafer = crispedricewafer,
    Hard = hard,
    Bar = bar,
    Sugar_Percentage = sugarpercent,
    Price_Percentage = pricepercent,
    Win_Percentage = winpercent
  )

# subset
candy_subsetdata <- candy_rawdata %>%
  select(
    Candy_Name, Chocolate, Fruity, Caramel, Peanut_Almond, Nougat, Crisped_Rice_Wafer, Hard, Bar, Sugar_Percentage, Price_Percentage, Win_Percentage
  )

head(candy_subsetdata)
```
### Sort Top 10 Popular Choice for Holloween
```{r}
top_10_candy <- candy_subsetdata %>%
  arrange(desc(Win_Percentage)) %>%
  head(10)

top_10_candy
```

### Create a Bar Plot
```{r}
ggplot(top_10_candy, aes(x = reorder(Candy_Name, Win_Percentage), y = Win_Percentage, fill = Candy_Name)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Top 10 Hallowwen Candies rated by Win Percentage",
    x = "Candy Name",
    y = "Win Percentage"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Conclusion 
The retailers and manufactureers should focus more on promoting chocolate and peanut-based candies during Halloween season based from graph above. Offering differient packs containing both with and without peanut chocolate may maximum the profit sales. 
